Self Referencing Sequences

### 2014-11-03 draft.tex
edited iterative process

We change focus in this section and consider how to generate self-referential sequences with an iterative process. We first look at the sequences generated by $\{1,2\}$.

One possible iterative process is as follows. Let the sequence $s^0$ be 2, and let the sequence $s^{(i)}$ for $i \geq 1$ be the unique sequence that starts with 2 and is such that its block lengths, read from left to right, reproduce the sequence $s^{(i-1)}$. This gives:
\begin{align*}
s^{(0)}&= 2\\
s^{(1)} &= 22\\
s^{(2)} &= 2211\\
s^{(3)}&= 221121\\
s^{(4)} &= 221121221\\
&\vdots
\end{align*}
Let $s^\infty$ be the limit of this process. 
By limit, we mean the limit with respect to finitely many terms $s^{(i)}_j \rightarrow s^\infty_j$ as $i \rightarrow \infty$. We claim that $s^\infty$ is the unique self-referential sequence $s(\{1,2\},2)$ generated by $\{1,2\}$ with starting number 2. In fact we can prove a more general result about this iterative process.

\begin{theorem}
Let $s^0 = a$ for some $a \in A$ where $a \neq 1$, and let $s^{(i)}$ for $i \geq 1$ be the unique sequence that starts with $a$ and is such that its block lengths, read from left to right, reproduce the sequence $s^{(i-1)}$. Then the limit of this process, $s^\infty$, is the sequence $s(A,a)$.
\end{theorem}

\begin{proof}
We prove this with induction. We show that if $s^{(i)}$ is a prefix of $s(A,a)$, then $s^{(i+1)}$ is a longer prefix of $s(A,a)$. The base case of the induction is satisfied, as $s^{(0)}$ is a prefix of $s(A,a)$.

@@

It turns out that we can be less restrictive about what our starting sequence is. Going back to the case of $\{1,2\}$, we can let $s^{(0)}$ be \emph{any} sequence of 1's and 2's that starts with a 2. For example, suppose $s^{(0)} = 2,1,1,1,2$. The rules for generating $s^{(i)}$ give:
\begin{align*}
s^{(0)}&= \mathbf{2},1,1,1,2\\
s^{(1)} &= \mathbf{2,2,1},2,1,2,2\\
s^{(2)} &= \mathbf{2,2,1,1,2,1},1,2,1,1,2,2\\
s^{(3)} &= \mathbf{2,2,1,1,2,1,2,2,1,2},1,1,2,1,2,2,1,1\\
&\vdots
\end{align*}
Here, we bold the numbers that form a prefix of $s(\{1,2\},2)$. We note that $s^\infty = s(A,a)$ because because $s^{(0)}$ contains a 2 at the beginning, which by the above theorem will produce $s(A,a)$ in the limit. We call the rest of each $s^{(i)}$ that is not a prefix of $s({1,2},2)$ the \emph{tail}.

We now consider two different forms of convergence. We know that $s^{(i)}$ in some sense approaches $s(A,a)$, as arbitrarily many elements at the beginning of the sequence will match up with the beginning of $s(A,a)$ as $i$ increases. However, if the tail of the sequence never entirely disappears, then there does not exist an $i$ where $s^{(i)}$ is exactly a prefix of $s(A,a)$. Thus, we formalize two types of convergence
\begin{definition} We say that a sequence $s^{i}$ is weakly convergent to $s(A,a)$ if $s^\infty = s(A,a)$. We say that the sequence $s^{i}$ is strongly convergent if $s^{(i)}$ is a prefix of $s(A,a)$ for all $i \geq k$ for some $k$.
\end{definition}
It is unclear what conditions must be placed on the starting sequence in order to ensure strong convergence, although based on numerical simulation we make the following conjecture.

\begin{conjecture} A sequence $s^i$ is strongly convergent to $s(\{1,2\}, 2)$ if the starting sequence begins with a $2$ and contains no more than 3 elements. A sequence $s^i$ is strongly convergent to $s(\{1,2\}, 1)$ if the starting sequence begins with a $1$ and contains between 2 and 3 elements. A sequence $s^i$ is weakly convergent but not strongly convergent to $s(\{1,2\}, a)$ if the starting sequence begins with an $a \in \{1,2,\}$ and contains 4 or more elements.
\end{conjecture}
We make this conjecture on the basis on computer simulations which executes the iterative process to at least 20 iterations. More iterations were not completed because after this point the simulations began to take impractically long amounts of time to complete.

@@

\noindent For the final draft,\\
we will improve the haikus\\
(and the whole paper).
\end{document}  
>>>>>>> 94de583213b09230c51ca57ed5c9897bcabc4666

>>>>

We change focus in this section and consider how to generate self-referential sequences with an iterative process. We first look at the sequences generated by the cycle $C = (1,2)$.

One possible iterative process is as follows. Let the sequence $s^0$ be 2, and let the sequence $s^{(i)}$ for $i \geq 1$ be the unique sequence that starts with 2 and is such that its block lengths, read from left to right, reproduce the sequence $s^{(i-1)}$. This gives:
\begin{align*}
s^{(0)}&= 2\\
s^{(1)} &= 22\\
s^{(2)} &= 2211\\
s^{(3)}&= 221121\\
s^{(4)} &= 221121221\\
&\vdots
\end{align*}
Let $s^\infty$ be the limit of this process. 
By limit, we mean the limit with respect to each term seperately, so we can define
\[ s^{(i)}_j \rightarrow s^\infty_j \]
as $i \rightarrow \infty$. We claim that $s^\infty$ is the unique self-referential sequence $s(C,2)$. In fact we can prove a more general result about this iterative process.

\begin{theorem}
Let $C$ be some generating cycle. Let $s^0 = a$ for some $a \in C$ where $a \neq 1$, and let $s^{(i)}$ for $i \geq 1$ be the unique sequence that starts with $a$ and is such that its block lengths, read from left to right, reproduce the sequence $s^{(i-1)}$. Then the limit of this process, $s^\infty$, is the sequence $s(C,a)$.
\end{theorem}

\begin{proof}
We prove this by induction on $i$. We show that if $s^{(i)}$ is a prefix of $s(C,a)$, then $s^{(i+1)}$ is a longer prefix of $s(C,a)$. The base case of the induction is satisfied, as $s^{(0)}$ is a prefix of $s(C,a)$.

@@

It turns out that we can be less restrictive about what our starting sequence, $s^{(0)}$, is. Going back to the case of $C = (1,2)$, we can let $s^{(0)}$ be \emph{any} sequence of 1's and 2's that starts with a 2. And still, the limiting sequence will be the self-referential sequence.
For example, suppose $s^{(0)} = 2,1,1,1,2$. The rules for generating $s^{(i)}$ give:
\begin{align*}
s^{(0)}&= \mathbf{2},1,1,1,2\\
s^{(1)} &= \mathbf{2,2,1},2,1,2,2\\
s^{(2)} &= \mathbf{2,2,1,1,2,1},1,2,1,1,2,2\\
s^{(3)} &= \mathbf{2,2,1,1,2,1,2,2,1,2},1,1,2,1,2,2,1,1\\
&\vdots
\end{align*}
Here, we bold the numbers that form a prefix of $s(C,2)$. We note that $s^\infty = s(C,a)$ because because $s^{(0)}$ contains a 2 at the beginning, which by the above theorem will produce $s(C,a)$ in the limit. We call the rest of each $s^{(i)}$ that is not a prefix of $s(C,2)$ the \emph{tail}.

We now consider two different forms of convergence. We know that $s^{(i)}$ in some sense approaches $s(C,a)$, as arbitrarily many elements at the beginning of the sequence will match up with the beginning of $s(C,a)$ as $i$ increases. However, if the tail of the sequence never entirely disappears, then there does not exist an $i$ where $s^{(i)}$ is exactly a prefix of $s(C,a)$. Thus, we formalize two types of convergence
\begin{definition} We say that a sequence $s^{i}$ is weakly convergent to $s(C,a)$ if $s^\infty = s(C,a)$. We say that the sequence $s^{i}$ is strongly convergent if $s^{(i)}$ is a prefix of $s(C,a)$ for all $i \geq k$ for some $k$.
\end{definition}
It is unclear what conditions must be placed on the starting sequence in order to ensure strong convergence, although based on numerical simulation we make the following conjecture.

\begin{conjecture} Let $C = (1,2)$. A sequence $s^{(i)}$ is strongly convergent to $s(C, 2)$ if the starting sequence begins with a $2$ and contains no more than 3 elements. A sequence $s^{(i)}$ is strongly convergent to $s(C, 1)$ if the starting sequence begins with a $1$ and contains between 2 and 3 elements. A sequence $s^{(i)}$ is weakly convergent but not strongly convergent to $s(C, a)$ if the starting sequence begins with an $a \in \{1,2\}$ and contains 4 or more elements.
\end{conjecture} 
We make this conjecture on the basis on computer simulations which executes the iterative process to at least 20 iterations. More iterations were not completed because after this point the simulations began to take impractically long amounts of time to complete.

@@

\noindent For the final draft,\\
we will improve the haikus\\
(and the whole paper).
\end{document}
